---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Methods

```{r setup}
library(tidyverse)
library(readxl)
library(modelsummary)

```



We describe our methods in this chapter.


Describe what data we have, and what it represents

Cleaning up the Data:

Storm data
Average data into 30 minute increments (both snowfall rate and storm severity index; include data/time & RWIS Site)
If both snowfall rate and storm severity index are 0, then get rid of it
Reassign storm severity index for <.25 as light, .25-1 as moderate and >1 as heavy

Speed Data
Average the data into 30 min increments (speed, Date/time, segment, Time_Period, weather type)

Table \@ref(tab:storm-table)

```{r ReadData} 
#can put extra options in here, such as a figure title in the chunk options

Storm <- read_xlsx("Data/Storm_Severity.xlsx")
Speed <- read_excel("Data/iPeMs.xlsx")

```

```{r storm-table}
#this is the table for the output data

datasummary_skim(Storm, title = "Distribution of Numeric Variables in Storm Data")
#shows mean, SD, min, median, max, histogram


```

```{r summaryofstormdata}

#the title is the caption
#put the comments above the figure being printed!
datasummary_skim(Storm %>%select("Storm Rating"=Adj_Rating, RWIS_Site), 
                 type="categorical", 
                 title = "Distribution of Categorial Variables in Data")


```

```{r speed-table}

datasummary_skim(Speed,
                 title = "Distirbution of Numeric Varaibles in Storm Data")

```

```{r summaryofspeeddata}

datasummary_skim(Speed %>%select(Segment, "Time of Day"=Time_Period, "Weather Type"=Date_Type),
                 type = "categorical",
                 title = "Distribution of Categorical Variables in Data")

```

##how to do a table without datasummary

|Threshold|Start Date|End Date|
|---------|----------|--------|
Can start a new row here
