# Applications

This is my attempt to figure out why my variables couldn't be called in later R chunks.

```{r,}
library(tidyverse)
library(readxl)
library(modelsummary)
library(lubridate)

Storm <- read_xlsx("Data/Storm_Severity.xlsx")
Speed <- read_xlsx("Data/iPeMs.xlsx")

storm_hour <- Storm %>%
  mutate(
    date = date(SampleTime),
    hour = hour(SampleTime)
  ) %>%
  group_by(RWIS_Site, date, hour) %>%
  summarise(
    s_intensity = mean(StormIntensityIndex, na.rm = TRUE),
    snowfall = mean(SnowfallRate, na.rm = TRUE)
  )

speed_hour <- Speed %>%  mutate(
    date = date(`Date/Time`),
    hour = hour(`Date/Time`)
  ) %>%
  group_by(Segment, date, hour) %>%
  summarise(
    Speed = mean(Speed, na.rm = TRUE),
    Speed_Limit = mean(Speed_Limit, na.rm = TRUE)
  )
```

The datasets were correlated together to find results. Traffic speed data were used to compare normal versus snow days to determine the effect of snow storms on traffic speeds. Storm severity was then compared to weather data to determine the effects of levels of severity on traffic speeds. The comparison of the three datasets showed the interaction between storm severity, traffic speed, and crash rates. 

The code below shows what would have been a regression line if the code would allow me to figure out why it wasn't calling out snowfall. I checked spelling and syntax, and tried to bring in the previous code, looked online, but with flying home today and other finals, this is the best I can do.

```{r}
ggplot(Storm, aes(x = snowfall, y = Speed, color = Speed)) +
  labs( x = "Snowfall Amount per Hour", y = "Traffic Speeds", title = "Regression of Snowfall v Traffic Speed") +
  geom_point() +
  geom_smooth(method = lm)
```

Here is the code that was used as a cross walk between the two datasets.
```{r, }
segment_xwalk <- tibble(
  Segment = c("W 9000 S","E 9000 S","Redwood","Redwood","Redwood","700 E","700 E","Foothill"),
  RWIS_Site = c("SR-85@DannonWay","I-15@6200S","I-15@6200S","I-15@500S","I-15@2400S","I-15@6200S","I-15@2400S","I-80@MouthofParleys")
)

segment_xwalk %>%
  left_join(storm_hour, by = "RWIS_Site") %>%
  left_join(speed_hour, by = c("Segment","date","hour"))
```

Let's try to perform a t-test on the data here. These t-tests were hoped to show the statistical evidence between 1) snowfall and speed and 2) storm intensity and speed. Again, this isn't working, but at this point, I don't know what else to do.

```{r}
t.test(snowfall, speed, paired=TRUE) 

t.test(s_intensity, speed, paired = TRUE)
```

I would do some analysis that would show histograms for different areas.

```{r}
hist(Segment, speed)
```

The last thing I hoped to do was to run an ANOVA test on segments and RWIS Sites to determine the statistical difference of storm intensity within the different routes that are paired with RWIS sites. For the same set up, I hoped to also do snowfall.

I know this isn't great, but it's the best I can do for now. Thank you for all your help.
